---
title: "Data Analysis - Treat"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
library(gt)
library(magrittr)
library(tidyverse)
library(glue)
library(lubridate)
```

```{r}
rm(list = ls())
dat_nst <- readRDS("rdsobs/dat_nst.RDS")
dat_arr <- readRDS("rdsobs/dat_arr.RDS")
```

### Actual Analysis

**Things that are being analyzed:**

-   Length of FU = Median, IQR

-   Number of follow-up biopsies(total per pateint) = Median, IQR.

-   Time between each patient's subsequent biopsies. = Median, IQR. *This includes all pre,post, and turp biopsies for the patient*

-   Number of upgrade events from GGG to GGG + 1. = Median, IQR *Does not discriminate if same patent had 2x upgrades in GGG*

-   Time between first treatment event, and the last diagnostic biopsy event. = Median, Min, Max

-   Type of treatment that was first received by a patient. = n, %

```{r}
ids <- unique(dat_nst$ID)
ntot <- unique(dat_nst$ID) %>% length()


# Lenght of follow-up(years), meadian (IQR)
lenfu <- dat_nst %>% 
  select(Dx.to.LastFU) %>% 
  summarise(
    med = median(Dx.to.LastFU, na.rm = TRUE),
    firstq = quantile(Dx.to.LastFU,  na.rm = TRUE)[2],
    thirdq = quantile(Dx.to.LastFU,  na.rm = TRUE)[4]
    #missing = sum(is.na(Dx.to.LastFU))
  ) 

# Number of follow-up biopsies
numfu <- dat_nst[,c("ID","nest.Biopsy")] %>% 
    unnest(nest.Biopsy) %>% 
    filter(!is.na(Biopsy.Dat) & !is.na(Biopsy.Dat.Orig)) %>%
    group_by(ID) %>% 
    count(Biopsy.Dat.Orig) %>%
    ungroup() %>% 
    #group_by(Biopsy.Dat.Orig) %>% 
    summarize(
      med = median(n, na.rm = TRUE),
      max = max(n, na.rm = TRUE),
      min = min(n, na.rm = TRUE)
    )
#Time between biopsies
timebetweenbiopsies <- dat_nst[,c("ID","nest.Biopsy")] %>% 
  unnest(nest.Biopsy) %>% 
  group_by(ID) %>%
  arrange(Biopsy.Dat.Orig, by_group = TRUE) %>% 
  mutate(Biopsy.TimeToLast = interval(lag(Biopsy.Dat), Biopsy.Dat) %>% as.numeric("years")) %>%
  ungroup() %>% 
  summarise(
    med = median(Biopsy.TimeToLast, na.rm = TRUE),
    firstq = quantile(Biopsy.TimeToLast,  na.rm = TRUE)[2],
    thirdq = quantile(Biopsy.TimeToLast,  na.rm = TRUE)[4]
    )

#Number of upgrades at follow up biopsy to different GGG
#NOTE!!!!! This calculation does not distinguish between the same patient who unerwent 2x upgrades. Please clarify with Adam if this is the expected behaviour that he wants.  
upgrade_gg <- dat_nst[,c("ID","nest.Biopsy")] %>% 
  unnest(nest.Biopsy) %>% 
  group_by(ID) %>%
  arrange(Biopsy.Dat.Orig, by_group = TRUE) %>%
  mutate(
    # Create new row of upgraded_biopsies
    Biopsy.Up = case_when(
      Biopsy.GGG - lag(Biopsy.GGG) >= 1 ~ as.character(Biopsy.GGG),
      is.na(Biopsy.GGG) == TRUE ~ "currna",
      is.na(lag(Biopsy.GGG)) == TRUE ~ "lastnaorbothna",
      length(ID) == 1 ~ "oneonly",
      (Biopsy.GGG - lag(Biopsy.GGG)) <= 0 ~ "noup",
      )
    ) %>% 
  mutate(
    #If there is an upgraded biopsy, then calculate the time between that and last biopsy
    Biopsy.UpTime = ifelse(!(Biopsy.Up %in% c("curnaa","lastnaorbothna","oneonly,","noup")),
      interval(lag(Biopsy.Dat), Biopsy.Dat) %>% as.numeric("years"),
      NA)
    ) %>% 
  ungroup() %>%
  group_by(Biopsy.Up) %>% 
  summarise(
    # Biopsy up is a number, so need to count the amount of biopsies that were upgraded.
    numb = length(Biopsy.Up),
    percent = (length(Biopsy.Up)/ntot)*100
  )  %>% 
  filter(Biopsy.Up %in% as.character(0:9))


#Time to first treatment
treat_time <- dat_arr %>%
  select(ID,Biopsy.Dat,Biopsy.Dat.Orig,Trt.Dat,Trt.Mod) %>% 
  group_by(ID) %>%
  #Make sure first treatment data, and the next biopsy is post-treatment
  # & (lead(Biopsy.Dat.Orig) == "biopsy_posttreatment")
  filter(Trt.Dat == min(Trt.Dat)) %>% 
  #drop_na() %>%
  summarise(
    timett = interval(Biopsy.Dat, Trt.Dat) %>% as.numeric("years"),
    Trt.Mod = first(Trt.Mod)
    ) %>% 
  ungroup() %>%
  group_by(Trt.Mod) %>% 
  summarise(
    med = median(timett, na.rm = TRUE),
    min = min(timett),
    max = max(timett))
  
#Treatment received percentage
treat_type <- dat_arr %>%
  select(ID,Biopsy.Dat,Biopsy.Dat.Orig,Trt.Dat,Trt.Mod) %>% 
  group_by(ID) %>%
  #Make sure first treatment data, and the next biopsy is post-treatment
  # & (lead(Biopsy.Dat.Orig) == "biopsy_posttreatment")
  filter(Trt.Dat == min(Trt.Dat)) %>% 
  group_by(Trt.Mod) %>%
  summarise(
    numb = n(),
    percent = n()/ntot*100
  )

```

### Fusing the data to make rough columns

```{r}
#Make a list of all the seperate tibbles. 
subtable_treat_names <- c("lenfu", "numfu" ,"timebetweenbiopsies" ,"upgrade_gg" ,"treat_time" ,"treat_type")
subtable_treatRAW <- list(lenfu, numfu ,timebetweenbiopsies ,upgrade_gg ,treat_time ,treat_type)
names(subtable_treatRAW) <- subtable_treat_names
subtable_treatRAW

#TABLES WITH SUBTABLES
with_subtable <- c("upgrade_gg" ,"treat_time" ,"treat_type")

subtable_treat <- subtable_treatRAW
subtable_treat <- map2(subtable_treat,subtable_treat_names,function(x,y){
  x %<>% mutate_if(is.numeric, ~ round(.,2))
  size <- ncol(x)
  if(size == 2){x %<>% mutate(nc_data = glue("{.[[size-1]]} ({.[[size]]})"))}
  if((size >= 3) & (y %in% with_subtable)){x %<>%  mutate(nc_data = glue("{.[[size-1]]} ({.[[size]]})"))}
  if((size >= 3) & !(y %in% with_subtable)){x %<>% mutate(nc_data = glue("{.[[size-2]]} ({.[[size-1]]} - {.[[size]]})"))}
  if(y %in% with_subtable){
    x %<>% rename(nc_subhead = 1)
    x %<>% mutate(nc_subhead = nc_subhead %>% 
                    as.character() %>% 
                    str_replace_all("_"," ") %>% 
                    str_to_title())
  }
  x %<>% select(contains("nc_"))
}) 

#Creating finished table
fintable_treat <- map2(subtable_treat,
                       names(subtable_treat),
                       function(x,y){add_column(x,nc_head = y,.before = 1)}) %>% bind_rows()


#Saving data
saveRDS(subtable_treatRAW, file = "rdsobs/subtable_treatRAW.RDS")
write.csv(fintable_treat,"output/fintable_treat.csv")
```

```{r}
#Finally drawing rough table with libgt
gttable_treat <- fintable_treat %>%
  gt(rowname_col = "nc_subhead",
     groupname_col = "nc_head") %>% 
  tab_header(md("**Treatment outcomes**"),
             subtitle = md("Edmonton cohort"))
  
gttable_treat

```
