---
title: "Data-Preparations"
output:
  html_document:
    df_print: paged
---

```{r message=FALSE, warning=FALSE}
require(tidyverse)
require(lubridate)

rm(list = ls()) #Clear env
```

### Importing the raw CSV file

Importing the raw CSV file and convert to tibble

*MD5 of orginal CSV : a2dd0181aa782319ca8b89ad0e52c501*

```{r}
#Import of the file
datRAW <- read.csv("data/21-Jan-2021_AS_258_Cohort.csv",
                header = TRUE, 
                stringsAsFactors = FALSE, 
                strip.white = TRUE,
                sep = ',',
                na.strings=c(""," ","NA"))

datRAW <- dplyr::as_tibble(datRAW)
glimpse(datRAW);datRAW
```

### Cleaning up

Cleaning up the column names, convert dates to data-time format using lubridate, and then removing the "years" from duration.

```{r}
dat <- datRAW

#New col names
newcols <- c("ID","AgeDx","FamHx","Eth","LastFU","LastFU.Orig","Dx.to.LastFU","Biopsy.Dat","Biopsy.Dat.Orig","Biopsy.PSA","BioClosest.Dat","BioClosest.PSA","Biopsy.Vol","Biopsy.GGG","Biopsy.PosCores","Trt.Dat","Trt.Mod")
colnames(dat) <- newcols

#Change the dates into date-type
dat <- dat %>% 
  mutate(across(c("LastFU", "Biopsy.Dat","BioClosest.Dat","Trt.Dat"),ymd))

#Remove the "years" in Dx.to.LastFU
dat$Dx.to.LastFU <- dat$Dx.to.LastFU %>% str_remove_all(" years") %>% as.double()


#Change all into simple case, and replace all spaces with _.
dat <- dat %>% 
  mutate(across(where(is.character), str_to_lower)) %>% 
  mutate(across(where(is.character), ~str_replace_all(.x," ","_")))


#Assign degree of relatedness - See https://medicine.utah.edu/surgery/research/images/sparc-graphics/degrees-of-relatedness.pdf
corRef <- read.csv("cor_reference.csv",
                header = TRUE, 
                stringsAsFactors = FALSE, 
                strip.white = TRUE,
                sep = ',',
                na.strings=c(""," ","NA"))

dat <- mutate(dat,FamHx = str_remove_all(FamHx,"_")) %>% 
  left_join(corRef,by = "FamHx") %>%
  relocate(Cor, .after = FamHx )
```

## Removing outliers

Removing obvious outliers

```{r}
#Removing outliers
overview <- dat %>%
  select(where(is.numeric) & !c(ID)) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(
    n = length(value),
    mean = mean(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    iqr = IQR(value, na.rm = TRUE)
  )

#Remove all negatives
dat[dat <= 0] <- NA


```


### Nesting

Creating nested tibbles via Purr to simplify data analysis.

```{r, message=FALSE}

#Function to split the data_frame, and nest the data.
squisher <- function(input,col_list,newcolname){
  temp <- input %>% 
    select(c(ID,col_list)) %>%
    fill(ID,.direction = "down") %>%
    filter(rowSums(is.na(.[,col_list])) != ncol(.[,col_list])) %>% #Remove rows that only have NA for data
    group_by(ID) %>% 
    nest() %>% 
    ungroup()
  colnames(temp) <- c("ID",newcolname)
  return(temp)
}

cols_Base <- c("ID","AgeDx","FamHx","Eth","LastFU","LastFU.Orig","Dx.to.LastFU")

cols_Biopsy <- c("Biopsy.Dat","Biopsy.Dat.Orig","Biopsy.PSA","BioClosest.Dat","BioClosest.PSA","Biopsy.Vol","Biopsy.GGG","Biopsy.PosCores")

cols_FamHx <- c("FamHx","Cor")

cols_Treat <- c("Trt.Dat","Trt.Mod")


dat_Base <- dat %>% 
  select(cols_Base) %>% 
  filter(is.na(ID) == FALSE)
dat_FamHx <- squisher(dat,cols_FamHx,"nest.FamHx")
dat_Biopsy <- squisher(dat,cols_Biopsy,"nest.Biopsy")
dat_Treat <- squisher(dat,cols_Treat,"nest.Treat")

#Joining main data with nested tibles
dat_nst <- dat_Base %>%
  full_join(dat_FamHx,by = "ID") %>%
  full_join(dat_Biopsy, by = "ID") %>%
  full_join(dat_Treat, by = "ID")
rm(dat_Base, dat_FamHx, dat_Biopsy, dat_Treat)

#Arrange in descending order according to ID
dat_nst <- dat_nst %>% arrange(ID)

glimpse(dat_nst);dat_nst

```


### Unique ID

Add a unique ID to each of the rows in the non-nested data so that they can be arranged without issue.
```{r}
#Add an unique ID for row easier reorganization in dat object
dat_arr <- dat %>% 
  mutate(Rec = 1:nrow(dat),.after = "ID") %>%
  fill(ID,.direction = "down")
```

### Save

Save into compressed RDS object

```{r}
saveRDS (dat_nst, file = "rdsobs/dat_nst.RDS")
saveRDS (dat_arr, file = "rdsobs/dat_arr.RDS")
saveRDS (newcols, file = "rdsobs/rdsnewcol.RDS")

```
